### 2022-01-12 src-latex/parts/10-introduction.tex

% General context + scope limitation
NER is a fundamental step of the information extraction pipeline when processing historical listing, like directories for instance.
OCRed text is not sufficient to build a high level, semantic view of the collection under study;
indeed, being able to properly tag text tokens unlocks the ability to relate entities,
and provide colleagues from digital humanities with useful databases.

% Problematic
NER requires training data, and we have little training data for historical documents.
OCR is also noisy.
How can we extract reliable elements in this context?

% Contributions
\begin{enumerate}
    \item review state of the art regarding NER for historical documents (section 2)
    \item dataset (section 3)
    \item list usable techniques (more to evangelize the DAR community) (section 4)
    \item compare their performance under various training set sizes (thus annotation costs) and OCR noise levels (thus OCR choices/luck/availability) (sec. 5 and 6)
    \item suggest smart way to leverage side knowledge like lexicons (FIXME iif this works well, maybe a conclusion instead)
\end{enumerate}

>>>>

% General context + scope limitation
The surge of historical documents digitized by cultural heritage institutions and made publicly available on the Web creates a vast knowledge source that can nourish numerous applications in the fields of social sciences, environmental studies or education.
Most of these documents are texts, either printed or handwritten.
To make the content of these documents searchable and usable by everyone, the texts engraved within them needs to be extracted through optical character recognition (OCR). 
However, OCRed texts are not sufficient to build a high level, semantic view of a collection of documents. A second step thus aims at recognizing in these texts the pieces of information most likely to be searched for by users, such as named entities (people, organizations, dates, places, etc.). Indeed, being able to properly tag text tokens unlocks the ability to relate entities, and provide colleagues from e.g. digital humanities with databases ready for exploitation.

Being active research topics, OCR and named entity recognition (NER) are still difficult tasks when applied to historical text documents.
OCR approaches used for modern documents are likely to struggle even on printed historical documents due the multiple factors related to text readability (low resolution scans, inconsistent printing rules, artifacts, show-through), document complexity (intricate and inconsistent page layout, use of ancient fonts \& special glyphs) and the variability inherent to the huge diversity of historical sources.
\bertrand{J'ai tenté un premier refactoring, pas plus techniquement précis ceci dit.}
\textcolor{green}{Phrases précedentes à reprendre avec les termes adéquats + ajouter une ref.}
On the other hand, NER approaches developed for modern texts suffer from the time gap between the contemporary entities and the entities mentioned in older texts and from the errors introduced into the text by the OCR step.\textcolor{green}{ajouter une ref}

In this article we focus on a peculiar category of historical printed books: the city trade directories of Paris from the XIX\textsuperscript{th} century (1801 - 1854).
They contain hundred-pages long lists of people, along with their activity and their address, so much valuable fine-grained knowledge to study the social dynamics of the city over time.
Despite containing roughly the same information, the diversity in layouts, information organization and digitizing quality is significant, which hardens makes OCR and NER quite challenging.

In a recent work to locate potentially polluted urban soils \cite{bell2020automated}, trade directories were leveraged to identify and locate all the gas stations in the city of Providence during the XX\textsuperscript{th} century.
The proposed information extraction pipeline extract texts with OCR, identify directories entries and recognizes business type and address within each entry.
In the same way, we aim at producing structured spatio-temporal data from the Parisian trade directories.
These directories originate from several publishers, their content is organized following different indexing methods (by name, by activity or by address), printed in various layouts, use different fonts.
We therefore investigate several state-of-the-art OCR and NER approaches independently and in combination to assess their usability on the directories corpus.
The contributions of this article are the following:
% Contributions
\begin{enumerate}
    \item review state of the art regarding NER for historical documents (section 2)
    \item dataset (section 3)
    \item list usable techniques (more to evangelize the DAR community) (section 4)
    \item compare their performance under various training set sizes (thus annotation costs) and OCR noise levels (thus OCR choices/luck/availability) (sec. 5 and 6)
    \item suggest smart way to leverage side knowledge like lexicons (FIXME iif this works well, maybe a conclusion instead)
\end{enumerate}

### 2022-01-12 src-latex/parts/40-ner-approaches.tex

% !TeX root = ../main-paper.tex
\section{NER Approaches}

\begin{itemize}
    \item regexp
    \item Spacy CNN with tok2vec
    \item CamemBERT finetuned with wikiner-fr + gold dataset (clean texts)
    \item CamemBERT finetuned with wikiner-fr + gold dataset (clean texts + pretrained on raw texts (Pero ? Kraken ? Both ?)
\end{itemize}

>>>>

% !TeX root = ../main-paper.tex
\section{NER Approaches}
Four approaches, one rules-based, three learning-based.

\paragraph{Regexp}
::TODO::

\paragraph{Spacy 3 transition-based NER.}
The \textit{of the shelf} NLP library SpaCY offers a named entity recognition tool based on an ad hoc architecture which has not been yet published but is explained by the developers on their website. Although the latest version of SpaCy (v3) leverages BERT models, we could not use them as no NER pipeline is available for the French language at the time of our experiments\footnote{https://spacy.io/models/fr}. The vanilla NER pipeline is two-fold. Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}. The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}. As words are processed sequentially, their vectors are concatenated with those in the last known entities to encode the nearby predicted semantics as a dynamic attention mechanism. The classification layer relies on a finite-state machine whose transition probabilities are learned using a multi-layer perceptron.
In 2018~\cite{won2018} compared SpaCy's vanilla NER performances on historical texts about places with several other software packages (Stanford-NER, Ner-Tagger, Edinburgh Geopoarser and Polyglot). Results showed SpaCy to be average with a F1-score of 0.57. The SpaCy developers claims an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.

\paragraph{CamemBERT \& CamemBERT+pretraining on historical directories}
::TODO::

\begin{itemize}
    \item regexp
    \item CamemBERT finetuned with wikiner-fr + gold dataset (clean texts)
    \item CamemBERT finetuned with wikiner-fr + gold dataset (clean texts + pretrained on raw texts (Pero ? Kraken ? Both ?)
\end{itemize}

### 2022-01-12 src-latex/parts/50-xp-setup.tex
info on camembert pretraining

\subsection{Metrics}
\begin{itemize}
    \item NER quality: Precision, Recall, F1 score (idea: stick with standard spacy/sklearn metrics) \textcolor{blue}{Dans la plupart des articles que j'ai lus, c'est toujours une F-mesure qui est utilisée et parfois la précision et le rappel sont ajoutés.}
    \item OCR accuracy: outputs from standard tools (beware of unknown glyphs: do we need any special care?) \textcolor{blue}{Pour l'instant, j'ai croisé: CER, WER, Levenshtein et Jaccard. J'ai encore un état de l'art à lire sur le sujet: il y aura probablement d'autres mesures mentionnées.}

@@

\subsection{Training procedure}
5-fold CV for all models.
\begin{itemize}
\item Spacy CNN : no limits on the number of epochs, instead use patience=1600.
\item CamemBERT : 3 epochs.
\end{itemize}

@@

\subsection{Implementation details}
we use spacy \mcite{spacy}
will we use FLAIR as well?

>>>>

\subsection{Metrics}
\bertrand{Move after section Cases studies? }
\begin{itemize}
    \item NER quality: Precision, Recall, F1 score (idea: stick with standard spacy/sklearn metrics) \textcolor{blue}{Dans la plupart des articles que j'ai lus, c'est toujours une F-mesure qui est utilisée et parfois la précision et le rappel sont ajoutés.}
    \item OCR accuracy: outputs from standard tools (beware of unknown glyphs: do we need any special care?) \textcolor{blue}{Pour l'instant, j'ai croisé: CER, WER, Levenshtein et Jaccard. J'ai encore un état de l'art à lire sur le sujet: il y aura probablement d'autres mesures mentionnées.}

@@

\subsection{Cases studies}
\subsubsection{Experiment 1: NER sensibility to the number of training samples}
The first experiment evaluates the performances of the three models on training sets of different sizes. To do so we split the gold reference into a training set, a development set and a test set. The training set is then gradually reduced in size while maintaining the relative frequency of directories within. The training and testing procedure is the same for the three models.

As the form an structure of entries varies across directories collections and through time, the models may learn to overfit on a subset of directories with specific features. To reduce the evaluation bias we start by leaving out 3 directories (1690 entries, ~20\%) from the gold reference to test each model on unseen directories. Then a stratified sampling based on the entry directory name is applied to the remaining set to create a training (6373 entries, ~71\% of the gold reference) and a development set (709 entries, ~8\%). This sampling procedure is a convenient way to shape both sets to reflect the diversity of directories within the gold reference.

To generate smaller training sets we start from the initial training set and attractively split it in half using the same stratified sampling strategy. We stop if there is only one entry left in a directory or if the current train set contains less than 30 entries. Applying this procedure to the initial training set produced 8 training sets containing 49, 99, 199, 398, 796, 1593, 3186 and 6373 entries.

% Move to metrics ?
All metrics are evaluated on the test set, yet the biased model performances on the development sets are add in the paper material as additional information.

\subsubsection{Experiment 2: named entity recognition in the context of noisy OCR data}

\begin{itemize}
\item At least 1 run per dataset, more if possible.
\item Only use the best model; should be CamemBERT+pretrained
\item Train on reference gold, evaluate on reference gold AND ocr-gold (2 datasets)
\item split train/dev/test the same way than experiment 1. Use the exact same splitting for datasets so the clean test set, pero test set and tesseract test sets contains the same entries.
\end{itemize}

%\begin{itemize}
%\item All : leave out 3 directories (~20\% of the dataset) as test data. Then apply stratified sampling based on the directory of each entry to split the remaining into a train set (~72\%) and a dev set (~8\%).
%\item Spacy CNN : No limit on epochs, instead use patience 1600.
%\item CamemBERT: 3 epochs.
%\item CamemBERT pretrained: 3 epochs, pretraining on MLM + NSP with ~20000 raw texts extracted with Pero-OCR.
%\end{itemize}

@@

\subsection{Implementation details}
We use Spacy\mcite{spacy} for the CNN implementation
and Huggingface for the BERT 
tok2vec(words embeddings + encoding) + attention layer  +  transition-based model.

\paragraph{Huggingface CamemBERT}
CamemBERT + linear classifier as decision layer

we use spacy  for the CNN implementation and Huggingface for BERT.

will we use FLAIR as well? $\rightarrow$ No
